A Detailed Analysis of 2020 U.S. Presidential Election: A Demographic Analysis Through Multilevel Regression and Post-Stratification

### 2024-03-15 paper/paper.qmd
typos, errors

---
title: "2020 U.S. Presidential Election Forecast"
subtitle: "How Multilevel Regression and Post-Stratification Unveil Voter Tendencies"
author: 
  - Jeongwoo Kim
  - Jiwon Choi
thanks: "Code and data are available at: https://github.com/Kjeongwoo99/2020_US_Election_Forecast"
date: today
date-format: long
abstract: "In this study, we forecasted the 2020 U.S. Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data. Our findings reveal how demographic factors such as age, gender, education, race, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump. The analysis highlights significant variations in voter support across different states and educational backgrounds, underscoring the complex interplay of demographics in shaping electoral outcomes. This research contributes to our understanding of the American electoral landscape, demonstrating the critical role of demographic diversity in determining the direction of political preferences and election results."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The U.S. Presidential Election is a pivotal world political issue given its impacts on social, economic and cultural dimension across the world. The 2020 Election was unique in that it was held during COVID-19, characterized by unprecedented voter turnout and divided political gap between the Republicans and the Democrats. Joe Biden was elected as the President of the United States after 4 years of the former President Donald Trump era. Our research uses statistical methodologies to analyze how demographics factors such as education, state, and gender influence voter behavior and resulted in the victory of the Democratic party. By utilizing multilevel regression with post-stratification (MRP) to a large-scale national survey data and census information, we explain the roles of gender, education, and geographic location in shaping voter preferences, with a special focus on the support for Donald Trump.

Understanding the electoral outcomes is crucial, not only for politicians but for the broader public and policymakers aiming to grasp the evolving landscape of American politics. Traditional analyses often fall short of capturing the relationships between voter demographics and electoral preferences. There are growing gaps between the groups in each education, gender and state level and this study responds to this need by employing MRP, a technique that allows accurate inferences about the voting behavior of diverse population segments.

Our findings reveal significant variations in voter support across different states and educational backgrounds and gender. The analysis points to a divergent pattern in voting behavior among different educational groups, challenging stereotypical narratives about educational attainment and party allegiance. We find that there is an increase in support for the Republicans from the voters with high education level in this election. We also find the gender gap between men and women where men are more supportive of Trump while women tend to favor Biden. Lastly, we discover the the strong preference for Biden from big cities in the west and the east coast while Trump's supporters are concentrated in the mid-west and suburb cities.

The importance of our study extends beyond academic circles, offering valuable perspectives for political strategists, journalists, and citizens alike. By providing a clearer picture of the demographic factors that influenced the 2020 election, our research aids in the development of more informed political strategies and fosters a deeper understanding of American democracy. Furthermore, our methodology showcases the potential of advanced statistical techniques to enhance the accuracy and depth of electoral forecasts.

The paper begins by introducing the broader context and motivation behind our study. It then discusses the specifics of the data sources and the variables that we considered important. This is followed by presentation of the data and discussion of our findings regarding education, gender and state in voter preference.

@@

We have used two datasets for this study. One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group [@citeSurveyData], conducted on October 3, 2019. Second is the census data from IPUMS America Census Service [@citeIPUMS], which is used as the post-stratification data for the survey data to adjust the weight.

@@

This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 [@citeSurveyData]. The sample is weighted to represent the U.S. adult population [@citeSurveyData]. Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code [@citeSurveyData]. According to the data, Male make up 48.3% while female make up 51.3% [@citeSurveyData]. 74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black [@citeSurveyData]. 20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ [@citeSurveyData]. On average, 5.1 percent declined immediately among those who are selected for the survey. 16.7 percent of the respondents did not complete the survey. Another 5.9 percent were categorized as speeding or straight-line which means they completed the survey in less than 6 minutes or selected the same response for every question in the three policy question batteries. Leaving these out leave 72.4 percent of the original sample for the analysis.

The Nationscape survey's strength lies in its methodological rigor - the effectiveness in collecting large samples from the U.S. citizen and its weighting strategy desinged to mirror the U.S. adult population by including weight factors such as age, gender, race and income and more. As they filter out inaccurate or missing data, it makes sure that the data collected are accurate and ensures data integrity. While other datasets such as the General Social Survey (GSS) and the American National Election Studies (ANES) are available, the Nationscape dataset's frequency (surveys collected every week) give it an advantage in analyzing electoral trends and shifts in real-time. Its' extensive sample size also justifies the choice of this dataset.

For our analysis, we decided to focus on five demographics: age, gender, education, race and state. Age is important because in general, voters tend to become more conservative as they get older. To account for the age difference, we divided the age group into four categories: 18-29, 30-49, 50-69 and 70+.

Gender is also an important category because in general, men tend be more conservative and women tend to be more liberal. Recently, gender issues are growing social issues and this may affect the election, hence we wanted to explore how this affects our model.

Education is also an interesting factor. In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans [@Harris]. However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls [@Harris]. Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so [@Harris]. We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.

@@

IPUMS ("Integrated Public Use Microdata Series") is a website that offers database of samples of the American population from the American Community Surveys of 2000-present. These samples provide rich qualitative information on the long-term changes in the population. We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research. The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates. It then uses stratified sampling where the U.S population is broken down into sub-groups and initial weights are assigned to each respondent.

@@

All figures showed in the data section illustrate comparisons between survey data and post-stratification data across different variables. First, the data comparison by state shows that the correspondence between the survey and post-stratification data for each state is quite good, further emphasize the overall reliability of the survey methodology in capturing diverse demographic characteristics. (see @fig-data-distribution-by-state) In other four figures, orange bars represent post-stratification data, while green bars signify survey data. The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values. Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories. Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin. (see @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education)

@@

@tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not. Apparently, more than half of the respondents, which is 55.30% choose not to vote for Donald Trump. Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party. The data used to create these table is from the Democracy Fund + UCLA Nationscape [@citeSurveyData]. We see that Donald Trump and his party Republican are not expected to win the popular vote before we implement the model.

@@

The key steps in MRP involve initially selecting a dataset for model development. In this case, we utilized survey data from the Voter Study Group [@citeSurveyData]. The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}. Following model creation, it is then applied to a broader dataset to estimate population characteristics. For our analysis, Census data from IPUMS [@citeIPUMS] served as this larger dataset.

To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group [@citeSurveyData] and applying post-stratification with Census Data [@citeIPUMS]. Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.

@@

\begin{equation}
log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{sex} +
\beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
\label{eq:logit}
\end{equation}

In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis. The variables chosen for this project are sex, age, race, education, and state. These were selected because sex, age, and race are proven to be reliable indicators of voting preferences. This decision is based on patterns such as certain states consistently favoring the Republican party, while others alternate between Democratic and Republican. Education was chosen over income because it provides a clearer picture of an individual’s background than income does.

Once the logistic regression model is developed, we'll use the `predict()` function in R [@citeR] to apply our model to Census data [@citeIPUMS], breaking down the dataset into categories based on sex, race, age group, education level, and state. This will give us the likelihood of individuals within each category voting for Donald Trump. These predictions allow us to analyze potential outcomes like the popular vote winner or the electoral college vote distribution.

@@

@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data [@citeSurveyData]. It also includes error bars, indicating the confidence interval for each coefficient estimate. In interpreting these coefficients, it's essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.

@@

stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
  mutate(support = ifelse(trump_predict >= 0.5, "Trump", "Biden"))

# make a plot of the result for state 
stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
                               color = support)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  labs(title = "Estimated Proportion of State Voting for Donal Trump",
       x = "Percentage", y= "State",
       color = "Support") +
  scale_x_continuous(labels = scales::percent) +
  theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values = c("Blue", "Red"))
```

Figure @fig-distribution-by-state shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms. Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state. Horizontal lines extending from the dots represent error bars for these estimates. The length of each line indicates the uncertainty associated with each estimate. For instance, we can see that this uncertainty lies between 55% to slightly higher than 80% for Trump in MT (Massachusetts). The dashed green line in the middle at the 50% mark represents the threshold for majority support. On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest to the lowest.

From @fig-distribution-by-state, it seems that the majority of the states support Biden. Only 7 states out of 51 have its point estimate greater than 50% for Trump. The horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans. Excluding these contesting states, however, our model suggests that only 3 states are definitely in favor of Trump whereas 35 states are definitely supporting Biden.

@@

@fig-distribution-by-educationlevel presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'. Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump. The horizontal lines extending to the left and right of each dot represent the error terms around the estimate, which reflect the uncertainty.

It shows that regardless of education level, the level of support for Trump lies below 40% The Republican party does not have the majority, including the error bars across all education levels. Voters with "High school or less" education level appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree. Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.

@@

## Analyzing Voter Preferences Across Education and Gender

@@

@@

The increasing propensity of older voters to lean towards the Democratic nominee, potentially driven by concerns over COVID-19 and healthcare, marks a significant shift in voting patterns. Similarly, the racial dynamics observed, with non-white demographics showing strong support for the Democratic party, reflect broader national conversations on race, immigration, and identity.

@@

We were able to estimate the winning states for each candidate with some success. From the data and analysis outlined above, we believe that our model suggests that Biden would have won both the popular vote and the electoral votes and become the president of the United States. This was partially right in that he did win both but with a smaller margin that we had expected. In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president [@CNN].

>>>>

---
title: "A Detailed Analysis of 2020 U.S. Presidential Election: A Demographic Analysis Through Multilevel Regression and Post-Stratification"
subtitle: "The Impact of Gender, Educatoin, Geographic location on Voter Preferences"
author: 
  - Jeongwoo Kim
  - Jiwon Choi
thanks: "Code and data are available at: https://github.com/Kjeongwoo99/2020_US_Election_Forecast"
date: today
date-format: long
abstract: "In this study, we forecasted the 2020 U.S. Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data. Our findings reveal how demographic factors such as gender, education, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump. The analysis highlights significant variations in voter support across gender, states and educational backgrounds and how they shape electoral outcomes. This research contributes to our understanding of the American electoral landscape, demonstrating the critical role of demographic diversity in determining the direction of political preferences and election results."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

The U.S. Presidential Election is a significant political issue given its impacts on social, economic and cultural dimensions across the world. The 2020 Election was unique in that it was held during COVID-19, characterized by unprecedented voter turnout and a divided political gap between the Republicans and the Democrats. Joe Biden was elected as the President of the United States after 4 years of the former President Donald Trump era. Our research uses statistical methodologies to analyze how demographic factors such as education, state, and gender influence voter behavior and result in the victory of the Democratic party. By utilizing multilevel regression with post-stratification (MRP) to large-scale national survey data and census information, we explain the roles of gender, education, and geographic location in shaping voter preferences, focusing on support for Donald Trump.

Understanding the electoral outcomes is crucial, not only for politicians but for the broader public and policymakers aiming to grasp the evolving landscape of American politics. Traditional analyses often fall short of capturing the relationships between voter demographics and electoral preferences. There are growing gaps between the groups in each education, gender, and state level and this study responds to this need by employing MRP. This technique allows accurate inferences about the voting behavior of diverse population segments. The estimand of our research is the true effect of demographic factors— gender, education, and geographic location—on voter preferences for Donald Trump in the 2020 U.S. Presidential Election. This effect represents the real-world influence that the demographic variables affect the likelihood of an individual voting for Donald Trump with all other variables held constant. It shows the estimated shift in voter preference towards Donald Trump to one unit change in each demographic factor. We aim to estimate this true effect as accurately as possible using available data.

Our findings reveal significant variations in voter support across different states and educational backgrounds and gender. The analysis points to a divergent pattern in voting behavior among different educational groups, challenging stereotypical narratives about educational attainment and party allegiance. We find that there is an increase in support for the Republicans from the voters with high education levels in this election. We also find the gender gap between men and women where men are more supportive of Trump while women tend to favor Biden. Lastly, we discover the the strong preference for Biden from big cities in the West and the East Coast while Trump's supporters are concentrated in the mid-west and suburban cities.

The importance of our study extends beyond academic, offering valuable insights for political strategists, journalists, and citizens. By providing a clearer picture of the demographic factors that influenced the 2020 election, our research aids in the development of more informed political strategies and fosters a deeper understanding of American democracy.

The paper begins by introducing the broader context and motivation behind our study. It then discusses the specifics of the data sources and the variables that we considered important. This is followed by the presentation of the data and a discussion of our findings regarding education, gender, and state in voter preference.

@@

We have used two datasets for this study. One is the U.S. election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group [@citeSurveyData], conducted on October 3, 2019. Second is the census data from IPUMS America Census Service [@citeIPUMS], which is used as the post-stratification data for the survey data to adjust the weight.

@@

This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from July 2019 to February 2021 [@citeSurveyData]. The sample is weighted to represent the U.S. adult population [@citeSurveyData]. Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code [@citeSurveyData]. According to the data, males make up 48.3% while females make up 51.3% [@citeSurveyData]. 74.2% of the respondents are White, 6.8% are Asian/Pacific, and 12% are Black [@citeSurveyData]. 20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, and 3.3% are 70+ [@citeSurveyData]. On average, 5.1% declined immediately among those who were selected for the survey. 16.7% of the respondents did not complete the survey. Another 5.9% were categorized as speeding or straight-line which means they completed the survey in less than 6 minutes or selected the same response for every question in the three policy question batteries [@citeSurveyData]. Leaving these out leave 72.4% of the original sample for the analysis.

The Nationscape survey's strength lies in its methodological rigor - the effectiveness of collecting large samples from U.S. citizens and its weighting strategy designed to mirror the U.S. adult population by including weight factors such as age, gender, race, and more. As they filter out inaccurate or missing data, it makes sure that the data collected are accurate and ensures data integrity. While other datasets such as the General Social Survey (GSS) and the American National Election Studies (ANES) are available, the Nationscape dataset's frequency (surveys collected every week) gives it an advantage in analyzing electoral trends and shifts in real-time. Its' extensive sample size also justifies the choice of this dataset.

For our analysis, we decided to choose five demographics: age, gender, education, race, and state and focus on gender, education and state. Age is important because in general, voters tend to become more conservative as they get older. To account for the age difference, we divided the age group into four categories: 18-29, 30-49, 50-69 and 70+.

Gender is also an important category because in general, men tend to be more conservative and women tend to be more liberal. Recently, gender issues are growing social issues and this may affect the election, hence we wanted to explore how this affects our model.

Education is also an interesting factor. In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans [@Harris]. However, there has been a switch in this trend as 61% of non-college white voters showed their support whereas just 45% of college-educated white voters did in the exit polls [@Harris].

@@

IPUMS ("Integrated Public Use Microdata Series") is a website that offers a database of samples of the American population from the American Community Surveys of 2000-present. These samples provide rich qualitative information on the long-term changes in the population. We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research to avoid any potential effects of COVID-19. The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates. It then uses stratified sampling where the U.S population is broken down into sub-groups and initial weights are assigned to each respondent.

@@

All figures showed in the data section illustrate comparisons between survey data and post-stratification data across different variables. First, the data comparison by state shows that the correspondence between the survey and post-stratification data for each state is quite accurate, further emphasizing the overall reliability of the survey methodology in capturing diverse demographic characteristics. (see @fig-data-distribution-by-state) In other four figures, orange bars represent post-stratification data, while green bars signify survey data. The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values. Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories. Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin. (see @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education)

@@

@tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not. More than half of the respondents, which is 55.30% choose not to vote for Donald Trump. Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting each party. The data used to create these table is from the Democracy Fund + UCLA Nationscape [@citeSurveyData]. We see that Donald Trump and his party Republican are not expected to win the popular vote before we implement the model.

@@

The key steps in MRP involve initially selecting a dataset for model development. In this case, we utilized survey data from the Voter Study Group [@citeSurveyData]. The next step is to construct a model with this smaller dataset; here, we employed a logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}. It is then followed by appling post-stratification to a broader dataset to estimate population characteristics. For our analysis, Census data from IPUMS [@citeIPUMS] served as this larger dataset. Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a willingness to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.

@@

\begin{equation}
log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{gender} +
\beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
\label{eq:logit}
\end{equation}

In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis. The variables chosen for this project are gender, age, race, education, and state. These were selected because gender, age, and race are proven to be reliable indicators of voting preferences. This decision is based on patterns such as certain states consistently favoring the Republican party, while others alternate between Democratic and Republican. Education was chosen over income because it provides a clearer picture of an individual’s background than income does.

Once the logistic regression model is developed, we'll use the `predict()` function in R [@citeR] to apply our model to Census data [@citeIPUMS], breaking down the dataset into categories based on gender, race, age group, education level, and state. This will give us the likelihood of individuals within each category voting for Donald Trump. These predictions allow us to analyze potential outcomes like the popular vote winner or the electoral college vote distribution.

@@

@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data [@citeSurveyData]. It also includes error bars, indicating the confidence interval for each coefficient estimate. In interpreting these coefficients, it is essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.

@@

stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
  mutate(support = ifelse(trump_predict >= 0.5, "Yes", "No"))

# a state plot of the result
stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
                               color = support)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  labs(title = "Estimated Proportion of State Voting for Donal Trump",
       x = "Percentage", y= "State",
       color = "Support Trump") +
  scale_x_continuous(labels = scales::percent) +
  theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values = c("Blue", "Red"))
```

@fig-distribution-by-state shows us the estimated of proportion of support for Trump or not by state using MRP with the inclusion of error terms. Each dot represents the point estimate of the proportion of support for Trump (red) or support for any other candidates including Joe Biden (blue) in each state. Horizontal lines extending from the dots represent error bars for these estimates. The length of each line indicates the uncertainty associated with each estimate. For instance, we can see that this uncertainty lies between 55% to slightly higher than 80% for Trump in MT (Massachusetts). The dashed green line in the middle at the 50% mark represents the threshold for majority support. On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest to the lowest.

From @fig-distribution-by-state, it seems that the majority of the states do not support Trump. Only 7 states out of 51 have a point estimate greater than 50% for Trump. The horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans. Excluding these contesting states, however, our model suggests that only 3 states are definitely in favor of Trump whereas 35 states are not considering supporting for Trump.

@@

@fig-distribution-by-educationlevel presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'. Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump. The horizontal lines extending to the left and right of each dot represent the error terms around the estimate, which reflect the uncertainty.

It shows that regardless of education level, the level of support for Trump lies below 40% The Republican party does not have the majority, including the error bars across all education levels. Voters with a "High school or less" education level appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree. Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we expected.

@@

## Analyzing Voter Preferences Across Education

@@

## Voter Preferences Across Gender

@@

@@

We were able to estimate the winning states for each candidate with some success. From the data and analysis outlined above, we believe that our model suggests that Biden would have won both the popular vote and the electoral votes and become the president of the United States. This was partially right in that he did win both but with a smaller margin that we had expected. In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president [@CNN]. however, + add why this happened
